---
title: "01_metadata_extraction"
author: "Arthur"
date: "28 février 2020"
output: html_document
---

This first script is used to extract and validate metadata from high resolution images. It is adapted to the datastructure of DigitalGlobe products, including Quickbird and Worlview-02, 03 & 04

```{r setup}
rm(list=ls()) # clean R environment
library(sf) # for spatial simple feature objects
library(raster) # for spatial raster data
library(tidyverse) # for tidy data handling 
library(ggplot2) # for easy data plotting
library(leaflet) # for interactive maps
library(mapview) # for better interactive maps 
```


# What data do we have?

The first step is to catalogue everything that we got and determine how to read it efficiently. The directory has to be changed to a directory containing all the files to explore.

```{r finding images}
imgs_dir = "//Glaciolab/homes/degranda/MFFP" # directory of all satellite images

#### extract date & type from the name ####
imgs = as.data.frame(dir(imgs_dir, recursive = T, pattern = ".TIF$")) # creates a dataframe of all image names 
imgs = imgs %>%
  mutate(date = lubridate::ymd_hms(str_match(imgs[,1], ".*(MUL|PAN)/(.*)(.-M|-P)2AS")[,3]))
imgs_MUL = imgs %>% 
  filter(str_detect(imgs[,1], "_MUL"))
imgs_PAN = imgs %>% 
  filter(str_detect(imgs[,1], "_PAN"))
```

## Multispectral images

Let's see what kind of data is contained in the MUL images.

```{r MUL formats, results= "hide"}
MUL_summary=tibble()
for(i in 1:length(imgs_MUL$value)){
  r = brick(paste0(imgs_dir,"/",imgs_MUL$value[i]))
  r_summary = cbind(print(res(r)[1]),print(nlayers(r)),print(crs(r)),print(imgs_MUL[i,]))
  MUL_summary = rbind(MUL_summary,r_summary)
}
colnames(MUL_summary) = c("res","nlayers","crs","img","date")
MUL_summary = as.tibble(MUL_summary)
```

```{r print MUL formats}
print(tbl_df(MUL_summary), n=Inf)
```

We seem to have the same CRS information for all archive data, but a longlat projection for the new acquisitions. 
  
The number of bands and their resolution is also variable from **4 to 8 bands**, going from **1.2 to 2.4 meters** in the archives, and 1.8e-5 degrees in the new acquisitions.  
  
The time of capture is always between 16:00 and 16:40 GMT, which means **11:00 to 11:40 local time**.  
  
The date ranges from **2009 to 2019**.  
  
Let's see how this translates in the data by visualising data from each group  
**(archive and new acquisitions will be treated separately)**

### wv3 MUL 8b 1.2m

```{r wv3 MUL 8b 1.2m}
r = brick(paste0(imgs_dir,"/",imgs_MUL$value[1]))
r[r==0] = NA # make 0s into NA
r = trim(r) # adjusts the extent
names(r) = c("b1","b2","b3","b4","b5","b6","b7","b8") # name the bands
plot(r) # display the data
```

It seems some of the bands have weird values (>255) and some linear artefacts (b3 and b5).  

### wv2 MUL 8b 2m

```{r wv2 MUL 8b 2m}
r = brick(paste0(imgs_dir,"/",imgs_MUL$value[6]))
r[r==0] = NA # make 0s into NA
r = trim(r) # adjusts the extent
names(r) = c("b1","b2","b3","b4","b5","b6","b7","b8") # name the bands
plot(r) # display the data
```

### qb2 MUL 4b 2.4m

```{r qb2 MUL 4b 2.4m}
r = brick(paste0(imgs_dir,"/",imgs_MUL$value[29]))
r[r==0] = NA # make 0s into NA
r = trim(r) # adjusts the extent
names(r) = c("b1","b2","b3","b4") # name the bands
plot(r) # display the data
```

### wv4 MUL 4b 1.2m

```{r wv4 MUL 4b 1.2m}
r = brick(paste0(imgs_dir,"/",imgs_MUL$value[30]))
r[r==0] = NA # make 0s into NA
r = trim(r) # adjusts the extent
names(r) = c("b1","b2","b3","b4") # name the bands
plot(r) # display the data
```

### wv3 MUL 8b lat/long 

```{r wv3 MUL 8b lat/long}
r = brick(paste0(imgs_dir,"/",imgs_MUL$value[36]))
#r[r==0] = NA # make 0s into NA
#r = trim(r) # adjusts the extent
names(r) = c("b1","b2","b3","b4","b5","b6","b7","b8") # name the bands
plot(r) # display the data
r2 = projectRaster(r, crs="+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
plot(r2)
r
r2
```