---
title: "classification"
author: "Arthur de Grandpré"
date: "30 juin 2020"
output: html_document
---

This script is to separate terrestrial and aquatic pixels prior to glint correction on images that need it.  

To generate our land mask, we need to classify our images into aquatic and terrestrial habitats.  

Since the data is so heavy, the process has to streamlined in a parallel workflow that limits the memory requirements and allow all cores to participate.  

The classificator training should be made on whole images since it has to represent the whole scene.  

The classification should be made on image tiles, since it is very heaving in terms of computing.

#1. generating the random.forest model

this requires :
 - the whole image (using previous outputs)  
 - a whole raster segmentation (using Orfeo ToolBox)  
 - a training set (using QGIS)  

## 1.1 images to correct

Not all images will require glint correction. This should be evaluated visually by considering wether or not it is believed that sun glint caused by waves can affect the capacity to detect submerged vegetation.

This can be done within any GIS software. We'll use QGIS3.

```{r years_to_fix, eval=T}
years = c(2011,2013,2017,2019) # years where glint was an issue
input_dir = "D:/Arthur/digitalglobe_archives/destripe/all/" # path to the de-stripped BOA images
imgs = dir(input_dir,full.names=T, pattern=".tif$")

glint_imgs = as.vector(NULL)

for(i in seq_along(years)){
glinty = imgs[grepl(as.character(years[i]),imgs)]
glint_imgs = append(glint_imgs,glinty)
}

glint_imgs # path to images to prepare for deglint
```

those are the 8 images that will be classified.

## 1.2 segmentation

### 1.2.1 feature extraction
For a better segmentation, additionnal features can be added to the image, such as texture, local statistics and edge extractions. All those layers can be used as additionnal information to delineate objects.

Let's define the feature extraction functions.

```{r localstat function, eval=T}
#####
feature.LocalStatisticExtraction <- function(
                           raster.in = "",
                           out.path  = "",
                           name      = "",
                           channel = "1",   #default 1, selected channel index in input image
                           radius   = "3"  #default 3, computational window radius
                           ){
  
# Set configuration      
conf <- paste("-RAM ",otb.ramlimit,
              "-in",raster.in,
              "-channel",channel,
              "-radius",radius,
              "-out",paste0(out.path,"/",name)
              )
  
  shell(paste("pushd ",otb.path,"&& otbcli_LocalStatisticExtraction ",conf))

write.table(x = conf,file = paste(out.path,"/",name,"_conf.txt",sep=""),row.names = F, col.names = F)
}

#####

feature.EdgeExtraction <- function(
                           raster.in = "",
                           out.path  = "",
                           name      = "",
                           channel = "1",   #default 1, selected channel index in input image
                           filter   = "gradient"  #default gradient, but can be changed for sobel or touzi
                           ){
  
# Set configuration      
conf <- paste("-RAM ",otb.ramlimit,
              "-in",raster.in,
              "-channel",channel,
              "-filter",filter,
              "-out",paste0(out.path,"/",name)
              )
  
  shell(paste("pushd ",otb.path,"&& otbcli_EdgeExtraction ",conf))

write.table(x = conf,file = paste(out.path,"/",name,"_conf.txt",sep=""),row.names = F, col.names = F)
}

#####

feature.HaralickTextureExtraction <- function(
                           raster.in = "",
                           out.path  = "",
                           name      = "",
                           channel = "1",   #default 1, selected channel index in input image
                           texture   = "simple",  #default simple, also available "advanced" or "higher"
                           parameters.min = "0", #input image minimum value
                           parameters.max = "1", #input image maximum value
                           parameters.xrad = "2", #xradius in pixels
                           parameters.yrad = "2", #yradisu in pixels
                           parameters.xoff = "1", #xoffset in pixels
                           parameters.yoff = "1", #yoffset in pixels
                           parameters.nbbin = "8" # bins per axis of histogram
                           
                           ){
  
# Set configuration      
conf <- paste("-RAM ",otb.ramlimit,
              "-in",raster.in,
              "-channel",channel,
              "-texture",texture,
              "-parameters.min",parameters.min,
              "-parameters.max",parameters.max,
              "-parameters.xrad",parameters.xrad,
              "-parameters.yrad",parameters.yrad,
              "-parameters.xoff",parameters.xoff,
              "-parameters.yoff",parameters.yoff,
              "-parameters.nbbin",parameters.nbbin,
              "-out",paste0(out.path,"/",name)
              )
  
  shell(paste("pushd ",otb.path,"&& otbcli_HaralickTextureExtraction ",conf))

write.table(x = conf,file = paste(out.path,"/",name,"_conf.txt",sep=""),row.names = F, col.names = F)
}
```

When calling the functions, sligthly different calls will be made based on wether the image has 4 or 8 bands in order to use the right NIR band.

```{r feature extractions, eval=F}
output_dir = "D:/Arthur/digitalglobe_archives/feature_extract/"

library(stringr)
library(raster)

otb.path  = "C:\\OTB-7.0.0-Win64\\bin"
otb.ramlimit = 2048 # does not seem to work without changing windows environment variable (not an issue)

for(i in seq_along(glint_imgs)){

  r = brick(glint_imgs[i])
  
  if(nlayers(r)==8){
feature.LocalStatisticExtraction(
                           raster.in = gsub("/","\\\\",glint_imgs[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"localstats"),
                           name      = paste0("ls_",str_sub(glint_imgs[i],-35)),
                           channel = "8",   #default 1, selected channel index in input image
                           radius   = "3"  #default 3, computational window radius
                           )
  
feature.EdgeExtraction(
                           raster.in = gsub("/","\\\\",glint_imgs[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"edge"),
                           name      = paste0("ee_",str_sub(glint_imgs[i],-35)),
                           channel = "8",   #default 1, selected channel index in input image
                           filter   = "gradient"  #default 3, computational window radius
                           )

feature.HaralickTextureExtraction(
                           raster.in = gsub("/","\\\\",glint_imgs[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"haralick"),
                           name      = paste0("ht_",str_sub(glint_imgs[i],-35)),
                           channel = "8",   #default 1, selected channel index in input image
                           texture   = "simple",  #default simple, also available "advanced" or "higher"
                           parameters.min = "0", #input image minimum value
                           parameters.max = "1", #input image maximum value
                           parameters.xrad = "2", #xradius in pixels
                           parameters.yrad = "2", #yradisu in pixels
                           parameters.xoff = "1", #xoffset in pixels
                           parameters.yoff = "1", #yoffset in pixels
                           parameters.nbbin = "8" # bins per axis of histogram
                           )
  } else {
feature.LocalStatisticExtraction(
                           raster.in = gsub("/","\\\\",glint_imgs[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"localstats"),
                           name      = paste0("ls_",str_sub(glint_imgs[i],-35)),
                           channel = "4",   #default 1, selected channel index in input image
                           radius   = "3"  #default 3, computational window radius
                           )
  
feature.EdgeExtraction(
                           raster.in = gsub("/","\\\\",glint_imgs[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"edge"),
                           name      = paste0("ee_",str_sub(glint_imgs[i],-35)),
                           channel = "4",   #default 1, selected channel index in input image
                           filter   = "gradient"  #default 3, computational window radius
                           )

feature.HaralickTextureExtraction(
                           raster.in = gsub("/","\\\\",glint_imgs[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"haralick"),
                           name      = paste0("ht_",str_sub(glint_imgs[i],-35)),
                           channel = "4",   #default 1, selected channel index in input image
                           texture   = "simple",  #default simple, also available "advanced" or "higher"
                           parameters.min = "0", #input image minimum value
                           parameters.max = "1", #input image maximum value
                           parameters.xrad = "2", #xradius in pixels
                           parameters.yrad = "2", #yradisu in pixels
                           parameters.xoff = "1", #xoffset in pixels
                           parameters.yoff = "1", #yoffset in pixels
                           parameters.nbbin = "8" # bins per axis of histogram
                           )
  }
  }
  
```

Let's add those features to the BOA image befor segmentation.

```{r, eval=F}
for(i in seq_along(glint_imgs)){
 
 r = brick(glint_imgs[i])
  
  features = c("localstats","edge","haralick")
  
  for(j in seq_along(features)){
  r = addLayer(r, brick(dir(paste0(output_dir,features[j]),full.names=TRUE, pattern=".tif$")[i]))
  }
  
  writeRaster(r, paste0(output_dir,str_sub(glint_imgs[i],-35)), overwrite=T)
}
```

visual validation

```{r, eval=F}
initial = dir(input_dir, full.names = T, pattern=".tif")
final = dir(output_dir, full.names = T, pattern=".tif")

plot(brick(initial[1]), main = "BOA image"); plot(brick(final[1]), main = "BOA + NIR features image")

```

Now that the features have been extracted, we can proceed to segmentation.

## 1.2.2 segmentation

Image segmentation is used to create polygons based on a raster. In this case, we want to use an algorythm able to process large images with a large amount of layers. For this application, we will use OTB's MeanShift segmentation, which can process rapidly very large amounts of multiband data.  
  
Those resulting polygons will then be used to classify the image based on object properties instead of pixel values.
  
Let's define the segmentation function call to OTB

```{r define segmentation function, eval=T}
meanshift.segm <- function(raster.in = "",
                           out.path  = "",
                           name      = "",
                           filter.meanshift.spatialr = "5",   #default 5
                           filter.meanshift.ranger   = "0.003",  #default 15
                           filter.meanshift.thres    = "0.001", #default 0.1
                           filter.meanshift.maxiter  = "100", #default 100
                           filter.meanshift.minsize  = "10"  #default 100
                           
                           ){
  
# Set configuration      
conf <- paste("-in",raster.in,"-filter meanshift","-filter.meanshift.spatialr",filter.meanshift.spatialr,
                "-filter.meanshift.ranger",filter.meanshift.ranger,"-filter.meanshift.thres",filter.meanshift.thres,
                "-filter.meanshift.maxiter",filter.meanshift.maxiter,"-filter.meanshift.minsize",filter.meanshift.minsize,
                "-mode vector","-mode.vector.out",paste0(out.path,"/",name,".shp"))
  
  shell(paste("pushd ",otb.path,"&& otbcli_Segmentation ",conf))

write.table(x = conf,file = paste(out.path,"/",name,"_conf.txt",sep=""),row.names = F, col.names = F)
}
```

Segmentation parameters will greatly affect the segmentation quality. If segmentation quality seems bad after visual validation, it is possible to change the parameters in the following chunk. The current values are based on simple tweaking and visual cues.

The output from this section is a simple shapefile

```{r directories for seg, eval=F}
input_dir = "D:/Arthur/digitalglobe_archives/feature_extract"
output_dir = "D:/Arthur/digitalglobe_archives/segmentation_water"
```

```{r segmentation, eval=F}
imgs = dir(input_dir,full.names=T, pattern=".tif$")

 for(i in seq_along(imgs)){
meanshift.segm(filter.meanshift.spatialr = "5",      # default 5
               filter.meanshift.ranger   = "0.003",  # default 15
               filter.meanshift.thres    = "0.001",  # default 0.1
               filter.meanshift.maxiter  = "100",    # default 100
               filter.meanshift.minsize  = "10",     # default 100
               raster.in = gsub("/","\\\\",imgs[i]),
               out.path  = gsub("/","\\\\",output_dir),
               name      = str_sub(imgs[i],-35))

  }
```

### visual validation

```{r, eval=F}
initial = dir(input_dir, full.names = T, pattern=".tif")
final = dir(output_dir, full.names = T, pattern=".shp")

plot(brick(initial[1]), main = "BOA image"); plot(readOGR(final[1]), main = "segments")
plot(brick(initial[1])[[2]], main = "BOA image, zoom B2 + segments", xlim=c(664500,665000),ylim=c(5111500,5112000)); plot(readOGR(final[1]), main = "segments",add=T, xlim=c(664500,665000),ylim=c(5111500,5112000))
```




#2. classification using the random.forest model  

this requires :  
 - tiles (using rgdal)  
 - tile segmentation (using Orfeo ToolBox)  
 - predictions (using the random.forest model)  
 - untiling and dissolving (using sf)
 
## 2.1 tiling

Let's define our tiling function so it can be run in parralel without hogging the memory.

```{r tiling_function, eval=T}
make_tiles <- function(j, tile.tbl, 
                          out.path.tif = out.path.tif,
                          source = input_rasters){
  
  out.tif = paste0(out.path.tif, paste0("/T_",j), tile.tbl[j,"ID"], ".tif")
  
  if(!file.exists(out.tif)){
    m <- readGDAL(input_rasters[i], offset=unlist(tile.tbl[j,c("offset.y","offset.x")]),
                 region.dim=unlist(tile.tbl[j,c("region.dim.y","region.dim.x")]),
                 output.dim=unlist(tile.tbl[j,c("region.dim.y","region.dim.x")]),
                 silent = TRUE)
    
    if(!all(is.na(m@data[,1]))){
      writeGDAL(m, out.tif, type="Float32", 
                options="COMPRESS=DEFLATE")
  }
  }
}

```

```{r write_tiles, eval=T}
library(rgdal)
library(GSIF)
input_rasters = glint_imgs
out.path.tif = ("D:/arthur/digitalglobe_archives/water_mask/tiles/tif/")

i=7

obj = GDALinfo(input_rasters[i])
tile.lst = getSpatialTiles(obj, block.x=500, return.SpatialPolygons=TRUE)
tile.tbl = getSpatialTiles(obj, block.x=500, return.SpatialPolygons=FALSE)

library(snowfall)

sfInit(parallel=TRUE, cpus=parallel::detectCores())
sfExport("make_tiles","tile.tbl","i","out.path.tif","input_rasters","tile.lst")
sfLibrary(rgdal)
sfLibrary(plyr)
sfLibrary(rgeos)
sfLibrary(gdalUtils)

out.lst = sfClusterApplyLB(1:nrow(tile.tbl),
                           function(x){make_tiles(x, tile.tbl ,out.path.tif, source = input_rasters)})

sfStop()

```

## 2.2 tile segmentation

## 2.2.1 feature extraction
```{r feature extractions, eval=T}
input_tiles = dir(out.path.tif, full.names=T, pattern=".tif$")
output_dir = "D:/Arthur/digitalglobe_archives/water_mask/tiles/tif/feature_extract/"

otb.path  = "C:\\OTB-7.0.0-Win64\\bin"
otb.ramlimit = 2048

library(raster)

for(i in seq_along(input_tiles)){

  r = brick(input_tiles[i])
  
  if(nlayers(r)==8){
feature.LocalStatisticExtraction(
                           raster.in = gsub("/","\\\\",input_tiles[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"localstats"),
                           name      = paste0("ls_",sub(".tif.*","",sub(".*tif/","",input_tiles[i]))),
                           channel = "8",   #default 1, selected channel index in input image
                           radius   = "3"  #default 3, computational window radius
                           )
  
feature.EdgeExtraction(
                           raster.in = gsub("/","\\\\",input_tiles[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"edge"),
                           name      = paste0("ee_",sub(".tif.*","",sub(".*tif/","",input_tiles[i]))),
                           channel = "8",   #default 1, selected channel index in input image
                           filter   = "gradient"  #default 3, computational window radius
                           )

feature.HaralickTextureExtraction(
                           raster.in = gsub("/","\\\\",input_tiles[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"haralick"),
                           name      = paste0("ht_",sub(".tif.*","",sub(".*tif/","",input_tiles[i]))),
                           channel = "8",   #default 1, selected channel index in input image
                           texture   = "simple",  #default simple, also available "advanced" or "higher"
                           parameters.min = "0", #input image minimum value
                           parameters.max = "1", #input image maximum value
                           parameters.xrad = "2", #xradius in pixels
                           parameters.yrad = "2", #yradisu in pixels
                           parameters.xoff = "1", #xoffset in pixels
                           parameters.yoff = "1", #yoffset in pixels
                           parameters.nbbin = "8" # bins per axis of histogram
                           )
  } else {
feature.LocalStatisticExtraction(
                           raster.in = gsub("/","\\\\",input_tiles[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"localstats"),
                           name      = paste0("ls_",sub(".tif.*","",sub(".*tif/","",input_tiles[i]))),
                           channel = "4",   #default 1, selected channel index in input image
                           radius   = "3"  #default 3, computational window radius
                           )
  
feature.EdgeExtraction(
                           raster.in = gsub("/","\\\\",input_tiles[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"edge"),
                           name      = paste0("ee_",sub(".tif.*","",sub(".*tif/","",input_tiles[i]))),
                           channel = "4",   #default 1, selected channel index in input image
                           filter   = "gradient"  #default 3, computational window radius
                           )

feature.HaralickTextureExtraction(
                           raster.in = gsub("/","\\\\",input_tiles[i]),
                           out.path  = paste0(gsub("/","\\\\",output_dir),"haralick"),
                           name      = paste0("ht_",sub(".tif.*","",sub(".*tif/","",input_tiles[i]))),
                           channel = "4",   #default 1, selected channel index in input image
                           texture   = "simple",  #default simple, also available "advanced" or "higher"
                           parameters.min = "0", #input image minimum value
                           parameters.max = "1", #input image maximum value
                           parameters.xrad = "2", #xradius in pixels
                           parameters.yrad = "2", #yradisu in pixels
                           parameters.xoff = "1", #xoffset in pixels
                           parameters.yoff = "1", #yoffset in pixels
                           parameters.nbbin = "8" # bins per axis of histogram
                           )
  }
  }
  
```

Let's add those features to the BOA image befor segmentation.  
first: define function for building stacks

# TROUBLESHOOT
issue : tile 74 has no data for haralick, also check tile 98

```{r, eval=F}
library(raster)
library(landscapetools)

i=190 # tile 74
r = brick(input_tiles[i])
features = c("localstats","edge","haralick")
j=1
r2 = addLayer(r, util_rescale(brick(dir(paste0(output_dir,features[j]),full.names=TRUE, pattern=".tif$")[i])))
j=2
r3 = addLayer(r2, util_rescale(brick(dir(paste0(output_dir,features[j]),full.names=TRUE, pattern=".tif$")[i])))
j=3
r4 = addLayer(r3, util_rescale(brick(dir(paste0(output_dir,features[j]),full.names=TRUE, pattern=".tif$")[i]))) 

plot(r[[1:8]])
plot(r2[[9:12]])
plot(r3[[13]])
plot(r4[[14:21]])

# error on haralicks

r = brick(dir(paste0(output_dir,features[j]),full.names=TRUE, pattern=".tif$")[i])
summary(r)
r2 = util_rescale(r)
?util_rescale()
plot(r)

# it is an issue with the rescale feature... maybe it isn't necessary
```


```{r eval=T}

add_features = function(i, input_tiles = input_tiles, output_dir = output_dir){
  r = brick(input_tiles[i])
  
  features = c("localstats","edge","haralick")
  
  for(j in seq_along(features)){
  # r = addLayer(r, util_rescale(brick(dir(paste0(output_dir,features[j]),full.names=TRUE, pattern=".tif$")[i])))
    r = addLayer(r, brick(dir(paste0(output_dir,features[j]),full.names=TRUE, pattern=".tif$")[i]))
  }
  
  writeRaster(r, paste0(output_dir,sub(".tif.*","",sub(".*tif/","",input_tiles[i])),".tif"), overwrite=T)
}
```

then run it in parallel using the snowfall package

```{r, eval=T}
library(landscapetools)
library(snowfall)
library(parallel)

sfInit(parallel=TRUE, cpus=parallel::detectCores()-1)
sfExport("input_tiles","output_dir","add_features")
sfLibrary(raster)
sfLibrary(landscapetools)

out.lst = sfClusterApplyLB(1:length(input_tiles),
                           function(x){add_features(x,input_tiles, output_dir)})

sfStop()
```

## 2.2.2 tile segmentation

Let's use the same function and parameters from previously to segment the tiles

```{r directories for seg2, eval=T}
input_dir = "D:/Arthur/digitalglobe_archives/water_mask/tiles/tif/feature_extract"
output_dir = "D:/Arthur/digitalglobe_archives/water_mask/tiles/tif/feature_extract/shps"
```

```{r segmentation2, eval=T}
imgs = dir(input_dir,full.names=T, pattern=".tif$")

 for(i in seq_along(imgs)){
   meanshift.segm(filter.meanshift.spatialr = "5",      # default 5
               filter.meanshift.ranger   = "0.003",  # default 15
               filter.meanshift.thres    = "0.001",  # default 0.1
               filter.meanshift.maxiter  = "100",    # default 100
               filter.meanshift.minsize  = "10",     # default 100
               raster.in = gsub("/","\\\\",imgs[i]),
               out.path  = gsub("/","\\\\",output_dir),
               name      = sub(".tif.*","",sub(".*feature_extract/","",imgs[i])))

 }
   
```

### visual validation

```{r, eval=F}
initial = dir(input_dir, full.names = T, pattern=".tif")
final = dir(output_dir, full.names = T, pattern=".shp")

plot(brick(initial[1]), main = "BOA image"); plot(readOGR(final[1]), main = "segments")
```

# 3. land masking  
## 3.1 train the classifier

Since random forest is a supervised method of classification, it requires training data. Training data must be manually produced using any GIS software, but here, for open source purposes, we select QGIS. Once a proper training set is built (enough samples of enough classes for every image), whole image classification can be performed.

```{r eval=T}
input_rasters = "D:/arthur/digitalglobe_archives/feature_extract/" #input BOA images with additionnal features
input_t_obj = "../data/training_sets" #location of training sets.
```

```{r zonalstats function, eval=T}
minx = function(x, na.rm=T){min(x)}
maxx = function(x, na.rm=T){max(x)}
meanx = function(x, na.rm=T){if(length(x)==1){return(x)}else{sum(x)/length(x)}}
sdx = function(x, na.rm=T){if(length(x)==1){return(0)}else{sd(x)}}


# zonal_summary = function(x, na.rm=T){
#   minx = order(as.numeric(x))[1]
#   maxx = order(as.numeric(x))[length(x)]
#   if(length(x)==1){
#     meanx = x[1]
#     sdx = 0
#   } else {
#     meanx = sum(x)/length(x)
#     sdx = sd(x)
#   }
# }

```


```{r, eval=T}
library(spatialEco)
library(randomForest)
rasters = dir(input_rasters, pattern=".tif$", full.names=T)
training_objects = dir(input_t_obj, pattern=".gpkg$", full.names=T)

i=7
r = brick(rasters[i])
# must be altered for 4 bands images
names(r) = c("C","B","G","Y","R","RE","N1","N","LS1","LS2","LS3","LS4","EE","HA1","HA2","HA3","HA4","HA5","HA6","HA7","HA8")

r = dropLayer(r, "LS1")
r = dropLayer(r, "LS2")

s = readOGR(training_objects[4])
s = s[,2:dim(s@data)[2]]
s = spTransform(s, crs(r))

# zonal_s = zonal.stats(x = s, y = r, stats = c("min","sd","mean","max"))
zonal_s = zonal.stats(x = s, y = r, stats = c("minx","maxx","meanx","sdx"))
s@data = cbind(s@data,zonal_s)

s2 = s@data[is.finite(rowSums(s@data[,2:77])),]

rf = randomForest(formula = class~.,
                  data = s2,
                  proximity = T,
                  ntree = 20000,
                  # mtry = 20,
                  # nodesize = 10,
                  importance = T)

plot(rf)
varImpPlot(rf)
importance = as.data.frame(importance(rf))
head(importance, 10)
rf

rm(s);rm(s2);rm(importance);rm(zonal_s)

```

## 3.2 apply the classifier

```{r, eval=T}
classify_tiles = function(j,
                          in.path.r = tifs,
                          in.path.shp = shps,
                          out.path.shp.class = output_shps){

#   in.path.r = tifs
#   in.path.shp = shps
#   out.path.shp.class = output_shps
# j=1
  out.shp = paste0(paste0(out.path.shp.class,sub(".tif.*","",sub(".*extract/","",tifs))[j],".shp"))  

r = raster::brick(in.path.r[j])
names(r) = c("C","B","G","Y","R","RE","N1","N","LS1","LS2","LS3","LS4","EE","HA1","HA2","HA3","HA4","HA5","HA6","HA7","HA8")

r = dropLayer(r, "LS1")
r = dropLayer(r, "LS2")

pc = rgdal::readOGR(in.path.shp[j])
pc = rgeos::gBuffer(pc, byid=T, width=0)
pc = raster::crop(pc,r)
pc = spTransform(pc, crs(r))

# pc
# plot(pc)

zonal_s = spatialEco::zonal.stats(x = pc, y = r, stats = c("minx","maxx","meanx","sdx"))
pc2=pc
pc2@data = cbind(pc@data,zonal_s)

pc2@data = pc2@data[is.finite(rowSums(pc2@data[,2:77])),]
pc2@data$class = "unknown"

pc2@data$class = predict(rf, pc2@data, type="class")
pc3 = pc
pc3@data = merge(pc@data,pc2@data,all=T)

rgdal::writeOGR(pc3,
                dsn=paste0(out.shp),
                driver="ESRI Shapefile",
                layer=paste0("TC_",j),
                overwrite_layer = T)
}

```

```{r classify_tiles, eval=T}

tifs = dir("D:/arthur/digitalglobe_archives/water_mask/tiles/tif/feature_extract/", full.names=T, recursive=F, pattern=".tif$")
shps = dir("D:/arthur/digitalglobe_archives/water_mask/tiles/tif/feature_extract/shps/", full.names=T, pattern=".shp$")

i=7

output_shps = "D:/arthur/digitalglobe_archives/water_mask/tiles/classified/"

library(snowfall)

sfInit(parallel=TRUE, cpus=parallel::detectCores()-1)
sfExport("tifs","shps","output_shps","classify_tiles","rf","minx","maxx","meanx","sdx")
sfLibrary(rgdal)
sfLibrary(raster)
sfLibrary(randomForest)
sfLibrary(spatialEco)
sfLibrary(rgeos)
sfLibrary(sp)

out.lst = sfClusterApplyLB(1:length(tifs),
                           function(x){classify_tiles(x, tifs, shps, output_shps)})

sfStop()
```

```{r, eval=T}
library(sf)
library(tidyverse)

cs = dir(output_shps, pattern=".shp$",full.names=T)

mosaic_shps = function(u,cs){
  s = st_read(cs[u])
  diss = s %>% 
    group_by(class) %>% 
    summarise() %>% 
    st_buffer(0)
}

library(doParallel)
library(foreach)

cl=parallel::makeCluster(detectCores()-1)
registerDoParallel(cl)

m = foreach(u=1:length(cs), .combine=rbind, .packages=c("sf","tidyverse")) %dopar% {
  mosaic_shps(u, cs)
}

stopCluster(cl)

plot(m)
st_write(m,
         "D:/Arthur/digitalglobe_archives/water_mask/landmask_2019_10_P001.gpkg",
         driver="GPKG",
         delete_dsn=T)
glint_imgs
```

### TROUBLESHOOT

there are some tiles that remain empty after classification
let's identify the bad tiles

```{r, eval=F}
library(sf)
tiles = SpatialPolygonsDataFrame(tile.lst, data=data.frame(c(1:length(tile.lst@polygons))))
tt = spTransform(tiles, CRS("+init=epsg:4326"))

tt2 = st_transform(m, CRS("+init=epsg:4326"))

centers <- data.frame(gCentroid(tt, byid = TRUE))
centers$ID = row.names(tt)
library(leaflet)
leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=tt2) %>% 
  addLabelOnlyMarkers(data=centers,
                      label=~ID,
                      lng=~x,
                      lat=~y,
                      labelOptions = labelOptions(noHide = TRUE,
                                                  direction = 'top',
                                                  textOnly = TRUE))

#some tiles to check : 368 for NA
```


```{r, eval=F}

tifs = dir("D:/arthur/digitalglobe_archives/water_mask/tiles/tif/feature_extract/", full.names=T, recursive=F, pattern=".tif$")
shps = dir("D:/arthur/digitalglobe_archives/water_mask/tiles/tif/feature_extract/shps/", full.names=T, pattern=".shp$")

i=1

output_shps = "D:/arthur/digitalglobe_archives/water_mask/tiles/classified/"


in.path.r = tifs
in.path.shp = shps
out.path.shp.class = output_shps

in.path.r
  j=168 #368
out.shp = paste0(paste0(out.path.shp.class,sub(".tif.*","",sub(".*extract/","",tifs))[j],".shp"))  

r = raster::brick(in.path.r[j])
names(r) = c("C","B","G","Y","R","RE","N1","N","LS1","LS2","LS3","LS4","EE","HA1","HA2","HA3","HA4","HA5","HA6","HA7","HA8")

r = dropLayer(r, "LS1")
r = dropLayer(r, "LS2")

pc = rgdal::readOGR(in.path.shp[j])
pc = rgeos::gBuffer(pc, byid=T, width=0)
pc = crop(pc,r)

summary(r)
plot(pc)
# test using exact_extract (seems to be the same)
# https://cran.r-project.org/web/packages/exactextractr/readme/README.html
install.packages("exactextractr")
library(exactextractr)


spc = st_read(in.path.shp[j]) %>% st_buffer(0)
zonal_s = exact_extract(r,spc, c("min","max","mean","stdev"))

rt = r
rt = brick(r)
rt[is.na(rt[])] <- 1000 
plot(rt[[8]]) # a lot of actual NAs




zonal_s = zonal.stats(x = pc, y = r, stats = c("minx","sdx","meanx","maxx"))
pc2=pc
pc2@data = cbind(pc@data,zonal_s)

summary(zonal_s)
dim(pc@data)
dim(pc2@data)

# pc2@data = pc2@data[is.finite(rowSums(pc2@data[,2:77])),]
pc2@data$class = "unknown"

plot(r[[1]])
hist(r[[1]])

pc2@data$class = predict(rf, pc2@data, type="class", na.action=na.pass)
pc3 = pc
pc3@data = merge(pc@data,pc2@data,all=T)

factpal <- colorFactor(topo.colors(5), pc3@data$class)

leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=spTransform(pc3,CRS=("+init=epsg:4326")),
              color =~factpal(class)) %>% 
  addLegend(pal=factpal,
            values=pc3@data$class)

rgdal::writeOGR(pc3,
                dsn=paste0(out.shp),
                driver="ESRI Shapefile",
                layer=paste0("TC_",j),
                overwrite_layer = T)

  s = st_read(cs[j])
  diss = s %>% 
    group_by(class) %>% 
    summarise() %>% 
    st_cast("POLYGON")

leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=st_transform(s, CRS("+init=epsg:4326")),
              color=~factpal(class))

### THE ISSUE IS WITH THE DISSOLVE METHOD
plot(s)
diss = s %>% 
    group_by(class) %>% 
    summarise() %>%
    st_buffer(0) #adding this resolves some polygons
    # st_cast("POLYGON") # the cast still wipes some polygons...

leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=st_transform(diss, CRS("+init=epsg:4326")),
              color=~factpal(class),stroke="transparent")

```

tile368 : lots of NAs, where do they come from?

```{r, eval=F}
v = brick("D:/Arthur/digitalglobe_archives/water_mask/tiles/tif/T_368.tif")
v[is.na(v)] <- 1000 
plot(v[[8]]) # a lot of actual NAs
test = v[[8]]
test[is.na(test)] = 1000
t2 = crop(test,r)
plot(t2)


v = brick("D:/Arthur/digitalglobe_archives/destripe/all/2011-06-03_MUL_WV02_2011-06-03_010623053070_01_P001.tif")
plot(v[[8]]) # a lot of actual NAs
test = v[[8]]
test[is.na(test)] = 1000
t2 = crop(test,r)
plot(t2)


v = brick("D:/Arthur/digitalglobe_archives/boa/MUL/2011-06-03_MUL_WV02_2011-06-03_010623053070_01_P001.tif")
plot(v[[8]])
test = v[[8]]
test[is.na(test)] = 1000
t2 = crop(test,r)
plot(t2)

v = brick("D:/Arthur/digitalglobe_archives/arc/MUL/mosaics/2011-06-03_010623053070_01_P001.TIF_TOA_MOSAIC.tif")
test = v[[8]]
test[is.na(test)] = 1000
t2 = crop(test,r)
plot(t2)

# FROM TOA TO BOA, NA appears
```

