---
title: "05_landmask"
author: "Arthur de Grandpré"
date: '2023-01-24'
output: html_document
---


# Intro

This is the fifth script of the Open HRRS W2 workflow for open source high resolution remote sensing of vegetation cover optically complex waters.

This script is used to perform a first classification to separate terrestrial pixels from aquatic pixels. This is an optional step if the image does not require deglinting, but it is critical if the image does require deglinting. Since deglinting is not fully implemented in the workflow yet, a simple landmask that keeps floating and emerging vegetation is acceptable. For a full deglinting of water pixels, a mask including all objects that could generate a NIR signature, such as dense and high vegetation canopy is necessary.  

This step requires manual work in the form of constructed training sets (multipolygon files, as shp or gpkg) built from QGIS or ArcGIS.  

It treats one image at a time, in order to allow for selection of the right training sets.

Multiple images of the same region and satellite acquisition should be able to be classified based on the same classifier training.

# 1. Setting up the R environment

## 1.1. Libraries

First, load the required libraries. They can be installed using the *install.packages("package.name", dependencies = T)* function. 

Some packages might require a valid installation of Rtools 4.0 or higher

```{r, message=F, warning=F}
rm(list=ls()) ; gc()

library(dplyr)
library(tidyr)
library(sf)
library(terra)
library(exactextractr)
library(foreach)
library(future)
library(doFuture)
library(future.apply)
library(ranger)

```

## 1.2. Inputs / Outputs

```{r}
#Parallel setup
registerDoFuture()
plan(multisession)

#OTB setup
otb.path  = "C:\\OTB-7.2.0-Win64\\bin" ### according to installation path and version (must be adjusted and use the double backslash!)

#input
input_destripe = dir("../data/work/destripe_output/",
                     full.names=T,
                     pattern=".tif$")

input_features = dir("../data/work/segmentation_output/features/",
                    full.names=T,
                    pattern=".tif$")
input_training = "../data/training_sets/land/" #location of training sets.
training_objects = dir(input_training, 
                       full.names=T,
                       pattern = ".gpkg$")
input_destripe
input_features
training_objects

# index of files to use
destripe_position = 1
feature_position = 1
training_position = 1

# output
model_output = normalizePath(paste0(input_training,"/models/"))
landmask_output = "../data/work/land_classification_output/"
water_output = paste0(landmask_output,"/water_output/")
```

create the directories

```{r}
for(i in c(model_output,
           landmask_output,
           water_output)){
  if(file.exists(i)){}else{
    dir.create(i)}
}

# file.copy(from=boa_input, to=deglint_output, recursive=F, overwrite=F) # so all files are still within the same folder, even if they are not all corrected.

```

## 1.3. Define functions

### Segmentation

```{r define segmentation function, eval=T}
otb_seg_file = function(raster.in = NULL,
                        out.path = NULL,
                        name = "",
                        filter.meanshift.spatialr = "5",   #default 5
                        filter.meanshift.ranger   = "0.05",  #default 15
                        filter.meanshift.thres    = "0.05", #default 0.1
                        filter.meanshift.maxiter  = "100", #default 100
                        filter.meanshift.minsize  = "1",  #default 100
                        path = "C:\\OTB-7.2.0-Win64\\bin",
                        ramlimit = 16000){
  
  
  # ## test chunk
  # raster.in = normalizePath(tar_read(resample)[1])
  # name = basename(raster.in)
  # out.path = gsub("/","\\\\",output)
  # filter.meanshift.spatialr = "15"   #default 5 # 5(too many objects) to 3 (too few objects) to 4 (unbalanced)
  # filter.meanshift.ranger   = "0.03"  #default 15 # 0.01 (too many objects) to 0.05 (too few objects) to 0.02
  # filter.meanshift.thres    = "0.2" #default 0.1
  # filter.meanshift.maxiter  = "100" #default 100
  # filter.meanshift.minsize  = "1"  #default 100
  # path = "C:\\OTB-7.1.0-Win64\\bin"
  # ramlimit = 16000
  ## test chunk \
  
  otb.path <<- path
  otb.ramlimit <<- ramlimit
  
  # Set configuration      
  conf <- paste("-in",raster.in,"-filter meanshift","-filter.meanshift.spatialr",filter.meanshift.spatialr,
                "-filter.meanshift.ranger",filter.meanshift.ranger,"-filter.meanshift.thres",filter.meanshift.thres,
                "-filter.meanshift.maxiter",filter.meanshift.maxiter,"-filter.meanshift.minsize",filter.meanshift.minsize,
                "-mode vector","-mode.vector.out",paste0(out.path,"/",name,".shp"))
  
  shell(paste("pushd ",otb.path,"&& otbcli_Segmentation ",conf))
  
  write.table(x = conf,file = paste(out.path,"/",name,"_conf.txt",sep=""),row.names = F, col.names = F)
}


otb_seg_dir = function(input_source = NULL,
                       output_path = NULL,
                       filter.meanshift.spatialr = "15",   #default 5 # 5(too many objects) to 3 (too few objects) to 4 (unbalanced)
                       filter.meanshift.ranger   = "0.03",  #default 15 # 0.01 (too many objects) to 0.05 (too few objects) to 0.02
                       filter.meanshift.thres    = "0.2", #default 0.1
                       filter.meanshift.maxiter  = "100", #default 100
                       filter.meanshift.minsize  = "1",  #default 100
                       path = "C:\\OTB-7.2.0-Win64\\bin",
                       ramlimit = 16000){
  
  ### TESTING CHUNK
# output = landmask_output
# input_source = dir(paste0(".",output),full.names=T,pattern=".tif$")
# output_path = paste0(".",output, "seg")
#  filter.meanshift.spatialr = spatialr   #default 5 # 5(too many objects) to 3 (too few objects) to 4 (unbalanced)
#  filter.meanshift.ranger   = ranger  #default 15 # 0.01 (too many objects) to 0.05 (too few objects) to 0.02
# filter.meanshift.thres    = thresh #default 0.1
# filter.meanshift.maxiter  = "100" #default 100
# filter.meanshift.minsize  = "1"  #default 100
# path = otb.path
# ramlimit = 1600

  ### END OF TESTING CHUNK
  
  if(dir.exists(output_path)==F){
    dir.create(output_path)
  }else{
    unlink(dir(output_path, full.names=T))
  }
  
  otb.path <<- path
  otb.ramlimit <<- ramlimit
  
  imgs = normalizePath(input_source)
  output = normalizePath(output_path)
  
  for(i in seq_along(imgs)){
    # i=1
      otb_seg_file(raster.in = gsub("/","\\\\",normalizePath(imgs[i])),
                 out.path = gsub("/","\\\\",normalizePath(output)),
                 name = basename(input_source)[i],
                 filter.meanshift.spatialr = filter.meanshift.spatialr,   #default 5
                 filter.meanshift.ranger   = filter.meanshift.ranger,  #default 15
                 filter.meanshift.thres    = filter.meanshift.thres, #default 0.1
                 filter.meanshift.maxiter  = filter.meanshift.maxiter, #default 100
                 filter.meanshift.minsize  = filter.meanshift.minsize,  #default 100
                 path = path
    )
  }
  
  return(dir(output,full.names=T, pattern=".shp$"))
  
}

```

### Train

```{r}
rf_train = function(input_segments = segs2,
                    input_rasters = ras,
                    out = output){
  ### test ###
# input_segments = training_objects[training_position]
# input_rasters = input_features[features_position]
# out = model_output
  ####
  
  s = st_read(input_segments)
  r = rast(input_rasters)
  
  names(r) = c("C","B","G","Y","R","RE",
               "N1","N","NH1","NH2","NH3","NH4","NH5","NH6","NH7","NH8",
               # "BH1","BH2","BH3","BH4","BH5","BH6","BH7","BH8",
               "NDVI","SAVI","EVI","NDAVI","WAVI")
  
  zonal_s = exact_extract(r,s, c("min","max","mean","stdev"), progress=F)
  
  s2 = cbind(s,zonal_s) %>% 
    drop_na()
  
  mx = ranger::ranger(as.factor(class) ~ ., data = st_drop_geometry(s2[,-1]))
  
  saveRDS(mx, paste0(out,"/",basename(input_rasters),".RDS"))
  
  return(mx)
}
```

### Predict

```{r}
predict_sav = function(in_model = NULL,
                       in_raster = NULL,
                       in_seg = NULL,
                       out_path = NULL,
                       name = NULL,
                       chunk_size = 1000){
  
# test ###

  # in_model = in_model
  # in_raster = in_raster[1]
  # in_seg = in_seg[1]
  # out_path = out_path
  # name = gsub(".tif","", basename(in_raster)[1])

####

  r = rast(in_raster)
  
  if(terra::nlyr(r)==16){
    names(r) = c("b1","b2","b3","b4","b5","b6","b7","b8","ha1","ha2","ha3","ha4","ha5","ha6","ha7","ha8")
  }else{
    names(r) = c("C","B","G","Y","R","RE",
                 "N1","N","NH1","NH2","NH3","NH4","NH5","NH6","NH7","NH8",
                 # "BH1","BH2","BH3","BH4","BH5","BH6","BH7","BH8",
                 "NDVI","SAVI","EVI","NDAVI","WAVI")
  }
    
  sx = crop2raster_loop(in_raster,
                   in_seg,
                   chunksize = 100000,
                   write = F)
  
  s = sx %>% 
    st_make_valid() %>% 
    dplyr::filter(st_is(.,"POLYGON")) %>% 
    dplyr::mutate(class = as.factor(NA))
  
  zx = exact_extract(r,s, c("min","max","mean","stdev"), progress=F, max_cells_in_memory = 100000)
  
    s2 = cbind(s,zx) %>%
      drop_na(4)
  
  # tryCatch({
  #   s2 = cbind(s,zx) %>%
  #     drop_na(min.b1)
  # }, error = function(e){
  #   s2 = cbind(s,zx) %>%
  #     drop_na(4)
  # }
  #   
  # )
  

  model = readRDS(in_model)
  
  sx2 = s2 %>%
    dplyr::select(-2) %>% 
    dplyr::mutate(class = ranger::predictions(stats::predict(model, st_drop_geometry(.)))) %>% 
    dplyr::group_by(class) %>% 
    dplyr::summarise() %>% 
    st_buffer(0) %>% 
    st_make_valid() %>% 
    st_write(dsn = paste0(out_path,"/",name,".gpkg"),
             driver = "GPKG",
             delete_dsn = T)

}


predict_sav_dir = function(in_model = NULL, 
                           in_raster = NULL,
                           in_seg = NULL,
                           out_path = NULL){
  #### test ###
# in_model = in_model[i]
# in_raster = dir(paste0(output,"har"), full.names=T, pattern=".tif$")
# in_seg = dir(paste0(output,"seg"),full.names=T,pattern=".shp")
# out_path = paste0(output, "pred")

  ####
  if(dir.exists(out_path)==FALSE){
    dir.create(out_path)
  }
  
  unlink(paste0(out_path,"/*"))
  
  plan(sequential)
  
  future_mapply(predict_sav,
                in_model = rep(in_model,length(in_raster)),
                in_raster = in_raster,
                in_seg = in_seg,
                out_path = rep(out_path,length(in_raster)),
                name = gsub(".tif","", basename(in_raster)),
                future.seed = TRUE)
  
  return(dir(out_path, pattern=".gpkg$", full.names=T))
  
}


crop2raster = function(raster,
                       vector,
                       output){
  
  r = rast(raster) %>% 
    terra::as.polygons(trunc = T,
                       dissolve = T,
                       values = T)
    st_as_sf() %>%
    st_make_valid()
  
  v = st_read(vector) %>%
    st_make_valid() %>%
    # terra::vect(.) %>% 
    # terra::crop(.,r) %>% 
    # terra::writeVector(.,paste0(output,"/",basename(raster)), filetype="GPKG")
    st_intersection(r) %>%
    st_write(paste0(output,"/",basename(raster),".gpkg"))
}

crop2raster_loop = function(raster,
                            vector,
                            output,
                            chunksize = 20000,
                            write = T){
  
  options("scipen"=10)
  
  r = rast(raster)
  rp = r %>%
    terra::as.polygons(trunc = T,
                       dissolve = T,
                       values = T)

  rp2 = st_as_sf(rp) %>% 
    st_make_valid()
  
  chunk_size = chunksize
  intersect_poly = rp2
  large_poly = vector
  
  for(i in 1:ceiling(as.numeric(st_layers(large_poly)[4])/chunk_size)){
    # i=279
  
    chunk=st_read(large_poly, query = paste0("SELECT * FROM \"", sf::st_layers(large_poly)[1] ,"\" LIMIT ",chunk_size," OFFSET ",ifelse(i==1,0,((i-1)*chunk_size-(i-1))))) %>% 
      st_make_valid(.) %>% 
      st_buffer(.,dist=0) %>% 
      st_transform(.,st_crs(rp2))
    
    if(i == 1){
      assign("to_fill", st_intersection(chunk,rp2)) 
    }
    
    if(i != 1){
      to_fill = bind_rows(to_fill, st_intersection(chunk,rp2))
    }
  }
  if(write==T){
    st_write(to_fill, paste0(output,"/",basename(raster),".gpkg"), driver = "GPKG", append=F)
  } else {return(to_fill)}
  
}


```

### Tile classif

```{r}
classif_tiles = function(in_haralick,
                         in_image,
                         in_model,
                         output,
                         spatialr = "15",
                         ranger = "0.03",
                         thresh = "0.2"){
  
  # ### test ###

# in_haralick = input_features[feature_position]
# in_image = input_destripe[destripe_position]
# in_model = dir(model_output, full.names=T, pattern=".RDS$")[1] # or select the right output from 2.1. Train
# output = landmask_output
# spatialr = "15"
# ranger = "0.03"
# thresh = "0.2"

  # ###
  
  if(!dir.exists(paste0(output,"/masks"))){
    dir.create(paste0(output,"/masks"))
  }
  
  foreach(i = 1:length(in_image)) %do% {
    # i=1
    if(dir.exists(output)){
      unlink(dir(output, full.names=T, ".tif"))
    }else{
      dir.create(output)
    }
    
    if(dir.exists(paste0(output,"seg"))){
      unlink(dir(paste0(output,"seg"), full.names=T))
    }else{
      dir.create(paste0(output,"seg"))
    }
    
    if(dir.exists(paste0(output,"har"))){
      unlink(dir(paste0(output,"har"), full.names=T))
    }else{
      dir.create(paste0(output,"har"))
    }
    
    if(dir.exists(paste0(output,"pred"))){
      unlink(dir(paste0(output,"pred"), full.names=T))
    }else{
      dir.create(paste0(output,"pred"))
    }
    
    # make image tiles (terra)
    # i=1
    r = terra::rast(in_image[i]) %>% 
      terra::mask(rast(in_image[i])[[1]])
    
    agg = r %>%
      aggregate(., fact = 500, fun = min)
    
    tiles = r %>% 
      makeTiles(.,
                agg,
                filename = paste0(output,"/_.tif"),
                na.rm=T)
    
    
    # make haralick tiles (terra)
    
    r = terra::rast(in_haralick[i]) %>% 
      terra::mask(rast(in_haralick[i])[[1]])
    
    agg = r %>%
      terra::aggregate(., fact = 500, fun = min)
    
    tiles = r %>% 
      terra::makeTiles(.,
                       agg,
                       filename =  paste0(output,"/har/_.tif"),
                       na.rm=T)
    
    # segment
    
    otb_seg_dir(input_source = dir(output,full.names=T,pattern=".tif$"),
                output_path = paste0(output, "seg"),
                filter.meanshift.spatialr = spatialr,   #default 5 # 5(too many objects) to 3 (too few objects) to 4 (unbalanced)
                filter.meanshift.ranger   = ranger,  #default 15 # 0.01 (too many objects) to 0.05 (too few objects) to 0.02
                filter.meanshift.thres    = thresh, #default 0.1
                filter.meanshift.maxiter  = "100", #default 100
                filter.meanshift.minsize  = "1",  #default 100
                path = otb.path,
                ramlimit = 16000)

    # predict

    predict_sav_dir(in_model = in_model[i],
                    in_raster = dir(paste0(output,"har"), full.names=T, pattern=".tif$"),
                    in_seg = dir(paste0(output,"seg"),full.names=T,pattern=".shp"),
                    out_path = paste0(output, "pred"))
    
    # merge
    ctiles = dir(paste0(output,"pred"),full.names=T,pattern=".gpkg")
    m = foreach(j = 1:length(ctiles), .combine = rbind) %do% {st_read(ctiles[j])}
    
    write = m %>%
      # filter(!str_starts(class, "l_")) %>% 
      st_write(.,driver="gpkg",paste0(output,"masks/",gsub(".tif",".gpkg",basename(in_image[i]))), delete_layer=T)
    
  }
  
  return(dir(paste0(output,"masks/"), full.names=T, pattern=".gpkg"))

}

plot(m)

```



# 2. Apply land masking

## 2.1. Train

```{r}
m = rf_train(input_segments = training_objects[training_position],
             input_rasters = input_features[feature_position],
             out = model_output
             )
```

## 2.2. Classify

```{r}
classif_tiles(
  in_haralick = input_features[feature_position],
  in_image = input_destripe[destripe_position],
  in_model = dir(model_output, full.names=T, pattern=".RDS$"),
  output = landmask_output,
  spatialr = "15",
  ranger = "0.03",
  thresh = "0.2"
              )

```

## 2.3. Mask

```{r}

mask = st_read(dir(paste0(landmask_output,"/masks"),pattern=".gpkg$",full.names=T)[destripe_position]) %>% 
  filter(str_detect(class, "^w")) %>% 
  vect()

image = rast(input_destripe[destripe_position])

water_only = image %>% 
  mask(mask) %>% 
  crop(mask) %>% 
  terra::writeRaster(paste0(water_output,"/",basename(input_destripe[destripe_position])))

```

