---
title: "06_deglint"
author: "Arthur de Grandpré"
date: "10/02/2021"
output: html_document
---

# Intro

This is an incomplete script of the Open HRRS W2 workflow for open source high resolution remote sensing of vegetation cover optically complex waters.  
  
This script comes after land masking and will perform empirical deglinting of water pixels, based on the Hochberg 2003 method. In order to perform correctly, all non glint NIR sources should be masked out during the previous step. If it is not the case, non glint NIR signal will be wrongly corrected.  
  
A manually delineated polygon of an area of optically deep water is required to run this script.  

While the results appear usable, it is unclear whether they should be implemented to help the classification or only later to limit errors in complexity estimation.

In this script, every image has to be treated independently since they require different training sets and inputs.

# 1. Setting up the R environment

First, load the required libraries. They can be installed using the *install.packages("package.name", dependencies = T)* function. The velox package requires the devtools library and a special function to download it from Github instead of CRAN.

## 1.1. Libraries
```{r, message=F, warning=F}
# for velox installation 
# library(devtools)
# install_github("hunzikp/velox")


otb.path  = "C:\\OTB-7.2.0-Win64\\bin"
library(raster)
library(tidyverse)
library(sf) 
library(sp)
library(rgdal) 
library(parallel)
library(doParallel)
library(foreach)
library(snowfall)
library(GSIF)
library(plyr) # used in snowfall env
library(rgeos) # used in snowfall env
library(gdalUtils) # used in snowfall env
library(spatialEco) # used in classification
library(randomForest) # used for classification
library(velox) # used for classification
```


## 1.2. Input/Outputs

```{r}
input_imgs = "../data/work/land_classification_output/"
input_deep = "../data/deep_zones/"
tdi_table = read.csv("../data/radiometric_calibration_parameters.csv")
  
output_dir = "../data/work/land_classification_output/deglint/"
```

create the directories 

```{r}
for(i in c(output_dir)){
  if(file.exists(i)){}else{
    dir.create(i)}
}

```


## 2. glint correction

Select images to fix and corresponding deep zone shapefiles

```{r}
imgs = dir(input_imgs, pattern=".tif$", full.names=T)
to_fix = c(1)
gi = imgs[to_fix]

dz = dir(input_deep, pattern=".shp$", full.names=T)
corresponding_shp = c(1)
dz = dz[corresponding_shp]
```

Method : Hochberg et al. 2003, Sea Surface Correction of High Spatial Resolution Ikonos Images to Improve Bottom Mapping in Near-Shore Environments. IEEE Transactions on geoscience and remote sensing.

```{r}
i=1

r = brick(gi[i])
if(nlayers(r)==8){
  r = brick(gi[i])[[1:8]]
  names(r) = c("C","B","G","Y","R","RE","N","N2")
} else {
  r = brick(gi[i])[[1:4]]
  names(r) = c("B","G","R","N")
}

d = st_read(dz[i])

tdi_img = tdi_table[str_detect(as.character(gi[i]), as.character(tdi_table$id_part)),]
tdi_img$bands = gsub(" ","",tdi_img$bands)
tdi_img = subset(tdi_img, tdi_img$bands != "P")
tdi_img$tdi = as.numeric(tdi_img$tdi)
tdis = unique(tdi_img$tdi)

if(length(tdis)==4){
    tdi_img[tdi_img$tdi==tdis[2],]$tdi = tdis[4]
  }

tdis = unique(tdi_img$tdi)
  
for(k in seq_along(tdis)){
  
  bands = subset(tdi_img$bands, tdi_img$tdi==tdis[k])
    
  rs = subset(r,tdi_img[tdi_img$tdi==tdis[k],]$bands)
  d = st_transform(d, crs(rs))
  rdz = readAll(crop(rs,d))
  
  max.coords = xyFromCell(rdz,which.max(rdz[[nlayers(rdz)]]))
  max.coords.pixel = SpatialPoints(max.coords)
  max.pixel = raster::extract(rdz,max.coords.pixel)
  
  min.coords = xyFromCell(rdz,which.min(rdz[[nlayers(rdz)]]))
  min.coords.pixel = SpatialPoints(t(as.data.frame(min.coords[1,])))
  min.pixel = raster::extract(rdz,min.coords.pixel)
  
  Lg_nir = max.pixel[[nlayers(rs)]] - min.pixel[[nlayers(rs)]]
  
  fg = (rs[[nlayers(rs)]] - min.pixel[[nlayers(rs)]]) / Lg_nir
  
  Lg = list()
  
  r2 = rs

  for(b in 1:nlayers(rs)){
    Lg[[b]] = max.pixel[[b]] - min.pixel[[b]]
    r2[[b]] = rs[[b]] - fg * Lg[[b]]
  }

  names(r2)=names(rs)
  
  if(k==1){r3=r2} else{r3 = addLayer(r3,r2)}
  
  if(k==length(tdis)){
    r3 = r3[[names(r)]]
    
    writeRaster(r3, paste0(output_dir,gsub("(.*)output/","",gi[i])), overwrite=T)
  } 
}
# }

plot(r3[[2]])
```

# OLD

```{r}
p_class = dir("D:/arthur/digitalglobe_archives/water_mask/",pattern=".gpkg$",full.names=T)
deep_zones = dir("D:/arthur/digitalglobe_archives/water_mask/deep_zones",full.names=T, pattern=".shp$")
tdi_table = read.csv("../data/radiometric_calibration_parameters.csv")
deglint_out = "D:/arthur/digitalglobe_archives/deglint_hochberg/hochberg_destripe/"

ydf = data.frame(1:8)
ydf = cbind(ydf, c(2011,2013,2013,2013,2013,2017,2019,2019))
colnames(ydf)=c("index","year")

for(i in 3:length(p_class)){
  
r = brick(glint_imgs[i])
cs = st_read(p_class[i])

if(ydf[i,2]==2011){
  rules = r2011
  }

if(ydf[i,2]==2013){
  rules = r2013
}

if(ydf[i,2]==2017){
  rules = r2017
}

if(ydf[i,2]==2019){
  rules = r2019
}

wm = cs %>% 
  filter(class %in% rules) %>% 
  st_union() %>% 
  st_cast("POLYGON")

r = mask(r,as_Spatial(wm))

# glint correction

tdi_img = tdi_table[str_detect(as.character(glint_imgs[i]), as.character(tdi_table$id_part)),]
tdi_img$bands = gsub(" ","",tdi_img$bands)
tdi_img = subset(tdi_img, tdi_img$bands != "P")
tdi_img$tdi = as.numeric(tdi_img$tdi)
tdis = unique(tdi_img$tdi)

if(length(tdis)==4){
    tdi_img[tdi_img$tdi==tdis[2],]$tdi = tdis[4]
  }

tdis = unique(tdi_img$tdi)

if(nlayers(r)==8){names(r) = c("C","B","G","Y","R","RE","N","N2")}
if(nlayers(r)==4){names(r) = c("B","G","R","N")}
  
dz = readOGR(deep_zones[i])
  
for(k in seq_along(tdis)){
  
  bands = subset(tdi_img$bands, tdi_img$tdi==tdis[k])
    
  rs = subset(r,tdi_img[tdi_img$tdi==tdis[k],]$bands)
  dz = spTransform(dz, crs(rs))
  rdz = readAll(crop(rs,dz))
  
  max.coords = xyFromCell(rdz,which.max(rdz[[nlayers(rdz)]]))
  max.coords.pixel = SpatialPoints(max.coords)
  max.pixel = raster::extract(rdz,max.coords.pixel)
  
  min.coords = xyFromCell(rdz,which.min(rdz[[nlayers(rdz)]]))
  min.coords.pixel = SpatialPoints(t(as.data.frame(min.coords[1,])))
  min.pixel = raster::extract(rdz,min.coords.pixel)
  
  Lg_nir = max.pixel[[nlayers(rs)]] - min.pixel[[nlayers(rs)]]
  
  fg = (rs[[nlayers(rs)]] - min.pixel[[nlayers(rs)]]) / Lg_nir
  
  Lg = list()
  
  r2 = rs

  for(b in 1:nlayers(rs)){
    Lg[[b]] = max.pixel[[b]] - min.pixel[[b]]
    r2[[b]] = rs[[b]] - fg * Lg[[b]]
  }

  names(r2)=names(rs)
  
  if(k==1){r3=r2} else{r3 = addLayer(r3,r2)}
  
  if(k==length(tdis)){
    r3 = r3[[names(r)]]
    writeRaster(r3, paste0(deglint_out,str_sub(glint_imgs[i],-55)), overwrite=T)
  } 
}
}

```

```{r}
input_dir = "D:/arthur/digitalglobe_archives/deglint_hochberg/hochberg_destripe/"
output_dir = "D:/arthur/digitalglobe_archives/ready_mosaics/"

dir(input_dir,full.names=T,pattern=".tif$")

for(i in 1:length(glint_imgs)){
  # i=1
  r1 = brick(dir(input_dir,full.names=T,pattern=".tif$")[i])
  r2 = brick(glint_imgs[i])

  for(j in 1:nlayers(r1)){
    if(max(getValues(r1[[j]]), na.rm=T) == min(getValues(r1[[j]]), na.rm=T)){
      r1[[j]]=r2[[j]]  
    }
  }

  # r3 = merge(r1,r2)
  r3 = mosaic(r1,r2, fun=min)
  
  
  if(nlayers(r1)==8){names(r3) = c("C","B","G","Y","R","RE","N","N2")}
  if(nlayers(r1)==4){names(r3) = c("B","G","R","N")}
  writeRaster(r3, paste0(output_dir,str_sub(glint_imgs[i],-55)), overwrite=T)
}
```

